---
title: "l0learn_challenge"
author: "stephens999"
date: "2018-12-08"
output: workflowr::wflow_html
---

## Introduction

The aim here is to illustrate L0learn's performance on a challenging example that we also use to challenge susie.
```{r}
library("L0Learn")
library("genlasso")
```

Here we simulate some data with two changepoints, very close together.
(Note: a more challenging example still might make these even closer together?)
```{r}
set.seed(1)
x = rnorm(100)
x[50:51]=x[50:51]+8
plot(x, col="gray")
```

Now we set up the design matrix $X$ to use a "step function" basis.
```{r}
n = length(x)
X = matrix(0,nrow=n,ncol=n-1)
for(j in 1:(n-1)){
  for(i in (j+1):n){
    X[i,j] = 1
  }
}
```

Now try L0Learn with L0 penalty. (Note: The CS algorithm is the faster
and less accurate one.)
```{r}
y.l0.CS = L0Learn.fit(X,x,penalty="L0",maxSuppSize = 100,autoLambda = FALSE,lambdaGrid = list(seq(0.01,0.001,length=100)),algorithm="CS") 


y.l0.CDPSI = L0Learn.fit(X,x,penalty="L0",maxSuppSize = 100,autoLambda = FALSE,lambdaGrid = list(seq(0.0076,0.0075,length=1000)),algorithm="CDPSI") 

plot(x,main="green=CDPSI; red=CS")
lines(predict(y.l0.CS,newx=X,lambda=y.l0.CS$lambda[[1]][min(which(y.l0.CS$suppSize[[1]]>1))]),col=2)
lines(predict(y.l0.CDPSI,newx=X,lambda=y.l0.CDPSI$lambda[[1]][min(which(y.l0.CDPSI$suppSize[[1]]>1))]),col=3)

```


```{r}
head(y.l0.CS$suppSize[[1]])
head(y.l0.CDPSI$suppSize[[1]])
```

Note that the more sophisticated CSPSI algorithm does not
find the correct solution here because it does not give 2 changepoints -
all solutions have at least 4 changepoints. I guess it
might be possible to solve this by playing with the penalty?



