---
title: "pelt"
author: "stephens999"
date: "2018-10-23"
output: workflowr::wflow_html
---

## Introduction

The idea here is to check into why the default method for `cpt.mean` 
function did not perfom well on some examples.
```{r}
library(changepoint)
```

```{r}
set.seed(51)
true_mean = rep(c(-0.2,0.1,1,-0.5,0.2,-0.5,0.1,-0.2),c(137,87,17,49,29,52,87,42))
genomdat = list(x = rnorm(500, sd=0.2) + true_mean, true_mean=true_mean)
```

The `cpt.mean` default does not find any changepoints:
```{r}
genomdat.cp = cpt.mean(genomdat$x,method="PELT")
plot(genomdat.cp)
```

Interestingly if we just multiply the data by 10 we find many changepoints.
```{r}
genomdat.cp = cpt.mean(10*genomdat$x,method="PELT")
plot(genomdat.cp)
```

And if we multiply by 100 we find many, many changepoints
```{r}
genomdat.cp = cpt.mean(100*genomdat$x,method="PELT")
plot(genomdat.cp)
```

  
I suspect that the cost function may implicitly assume residual variance is 1; I opened an issue on the `checkpoint` github page.

