---
title: "susie-paper"
author: "stephens999"
date: "2018-11-12"
output: workflowr::wflow_html
---

## Introduction

```{r}
library("susieR")
library("DNAcopy")
library("changepoint")
```


## Simulated example from DNAcopy

This is the simulated example from `?DNAcopy::segment` (note we changed
this into a single problem (single chromosome) whereas the example was originally split into two chromosomes).
```{r}
set.seed(51)
genomdat <- rnorm(500, sd=0.2) +
rep(c(-0.2,0.1,1,-0.5,0.2,-0.5,0.1,-0.2),c(137,87,17,49,29,52,87,42))
chrom <- rep(1,500)
maploc <- c(1:500)
test2 <- DNAcopy::segment(CNA(genomdat, chrom, maploc))
plot(test2)
```

```{r "susie-change-point-example"}
genomdat.s = susieR::susie_trendfilter(genomdat,0)
plot_cs = function(s){
  CS = s$sets$cs
  for(i in 1:length(CS)){
    rect(min(CS[[i]]),-5,max(CS[[i]])+1,5,col = rgb(1,0,0,alpha=0.5),border=NA)
  }
}

plot(genomdat,col="gray",ylab="y",xlab="t")
segments(x0=test2$output$loc.start,x1=test2$output$loc.end,y0=test2$output$seg.mean,col="blue",lwd=2)
abline(v=cumsum(c(137,87,17,49,29,52,87))) # changepoint locations
plot_cs(genomdat.s)
```


Report CSs and compare with true locations
```{r}
genomdat.s$sets$cs
cumsum(c(137,87,17,49,29,52,87))
```

## Convergence issue

We set up a simple simulated example with two change points very close together.
```{r}
set.seed(1)
x = rnorm(100)
x[50:51]=x[50:51]+8
x.s = susie_trendfilter(x,0,L=2,estimate_prior_variance=TRUE)
plot(x,col="gray",ylab="y",xlab="t")
lines(predict(x.s),type="s")
susie_get_objective(x.s)
```

Now try initializing susie at the "truth"
```{r}
s0 = susie_init_coef(c(49,51),coef_value = c(8,-8),p=100)
x.s2 = susie_trendfilter(x,0,s_init=s0,estimate_prior_variance=TRUE)
plot(x,col="gray",ylab="y",xlab="t")
lines(predict(x.s2),col=2,type="s")
susie_get_objective(x.s2)
```

```{r "susie-converge-problem"}
plot(x,col="gray",ylab="y",xlab="t")
lines(predict(x.s),type="s")
lines(predict(x.s2),col=2,type="s")
```

