---
title: "Trend filtering"
author: "Matthew Stephens"
date: "April 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here I'm investigating the strategy of 3-stage fit:
i) run with small residual variance
ii) estimate residual variance
iii) estimate prior variance

I'm interested in whether it works in settings both
where truth is sparse and not sparse.

Because of the strong correlations among the Xs trend filtering seems
a challenging test case.

Simulate some data and perform trend filtering.
```{r}
library("susieR")
set.seed(1)
n=1000
delta = matrix(1,nrow=n,ncol=n)
for(i in 1:(n-1)){
  for(j in (i+1):n){
    delta[i,j] = -1
  }
}

beta = c(rep(0,100),rep(1,100),rep(3,100),rep(-2,100),rep(0,600))
y = beta + rnorm(n)
delta[,2:1000] = scale(delta[,2:1000])
s0.1 = susie(delta,y,L=10,residual_variance=0.01,estimate_residual_variance = FALSE)
s0.2 = susie(delta,y,s_init = s0.1,estimate_residual_variance = TRUE)
s0.3 = susie(delta,y,s_init = s0.2, estimate_prior_variance = TRUE)
```

Plot results: the truth is black, and susie estimates are other colors.
```{r}
plot(y)
lines(predict(s0.1),col=4,lwd=3)
lines(predict(s0.2),col=5,lwd=3)
lines(predict(s0.3),col=2,lwd=3)
lines(beta,col=1,lwd=3)
s0.3$sa2
```

so we see that the final fit does not use all 10 effects.

Try something harder where the beta increases linearly (so not really sparse!):
```{r}
set.seed(1)
beta = seq(0,2,length=1000)
y = beta + rnorm(n)
s1.1 = susie(delta,y,L=10,residual_variance=0.01,estimate_residual_variance = FALSE)
s1.2 = susie(delta,y,s_init = s1.1,estimate_residual_variance = TRUE)
s1.3 = susie(delta,y,s_init = s1.2, estimate_prior_variance = TRUE)
s1.3$sa2
plot(y)
lines(predict(s1.3),col=2,lwd=3)
lines(predict(s1.2),col=3,lwd=3)
lines(predict(s1.1),col=4,lwd=3)
lines(beta,col=1,lwd=3)
```
